---
title: 'Homework 1: Fun with Climate Trends!'
author: "Peyton Moore"
date: "4/14/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(lubridate)
library(ggplot2)
library(tidyverse)
library(janitor)
library(chron)
library(lubridate)
library(tsibble)
library(feasts)
library(slider)
```




### Introduction

- description of location choosen: Cape Florida, FL 
- description of data used : https://www.ncdc.noaa.gov/cdo-web/confirmation

```{r, reading in data}


florida <- read_csv("florida.csv") %>% 
  clean_names() %>% 
  mutate(date = as_date(date), 
         year = year(date), 
         month = month(date), 
         day = day(date)) %>% 
  select(date, year, month, day, prcp, tmax, tmin)
  
```



### Trend Analysis Results 

Assess whether climate at your location has been changing with trend analysis
- using 2 different metrics for averages
- using 2 different metrics for extremes


```{r daily plots}
ggplot(data = florida, 
       aes(x = date, y = tmax))+
  geom_line()

ggplot(data = florida, 
       aes(x = date, y = tmin))+
  geom_line()

ggplot(data = florida, 
       aes(x = date, y = prcp))+
  geom_line()


```


```{r, season plots }

florida_ts <- florida %>% 
  as_tsibble(key = NULL, 
             index = date) %>%
  fill_gaps() %>% 
  filter_index("1999"~"2020")

gg_season(data =florida_ts, aes(y = tmax))

```



### Discussion 
Why are these metrics likely to be relevant, given the climate change impacts you expect at this location?

Use some good old T, rank-sum, or other statistical tests,


### Citations 



