---
title: 'Homework 1: Fun with Climate Trends!'
author: "Peyton Moore & Zoe Durken-Salm"
date: "4/14/2021"
output: 
  html_document:
    code_folding: hide
    theme: cosmo
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(lubridate)
library(ggplot2)
library(tidyverse)
library(janitor)
library(chron)
library(lubridate)
library(tsibble)
library(feasts)
library(slider)
library(tmap)
library(leaflet)
library(leaflet.esri)
library(sf)
library(here)
```




### Introduction

Here we will investigating changing climatic trends from 1999-2000 in Cape Florida, FL. - high risk 
- miami dade county 
- cite paper everglades 

```{r, cape florida map, include = FALSE}
cape <- data.frame(
  site = c("Cape Florida"), 
  lat = c(25.6683433), 
  long = c(-80.1573228)
)

florida_map <- st_read(here::here("data", "florida_counties"), layer = "united_states_florida_administrative_boundaries_level6_counties_polygon")


tmap_mode("view")

counties_map <- tm_shape(florida_map)+
  tm_fill("name", 
          alpha = 0.35, 
          legend.show = FALSE)

cape_map <- tmap_leaflet(counties_map)
```

```{r, final map, include = TRUE}
cape_map %>% 
  leaflet::addCircleMarkers(data = cape, 
                                               lng = ~long, 
                                               lat = ~lat, 
                                               label = c("Cape Florida"), 
                                               color = "orange", 
                                               labelOptions = labelOptions(noHide = T, 
                                               textsize = "15px")) %>% 
  addTiles() %>% 
  addEsriBasemapLayer(esriBasemapLayers$Imagery)

```




- description of location choosen: Cape Florida, FL 
- description of data used : https://www.ncdc.noaa.gov/cdo-web/confirmation

```{r, reading in data}


florida <- read_csv("florida.csv") %>% 
  clean_names() %>% 
  mutate(date = as_date(date), 
         year = year(date), 
         month = month(date), 
         day = day(date)) %>% 
  select(date, year, month, day, prcp, tmax, tmin)
  
```



### Trend Analysis Results 

Assess whether climate at your location has been changing with trend analysis
- using 2 different metrics for averages
- using 2 different metrics for extremes





```{r, tsibble, include = FALSE }

florida_ts <- florida %>% 
  as_tsibble(key = NULL, 
              index = date)



florida_month <- florida_ts %>%  
  index_by(yr_mo = ~yearmonth(.)) %>% 
  summarize(monthly_mean_temp = mean(tmax, na.rm = TRUE))


```

```{r, annual, include = FALSE}
fl_annual <- florida %>% 
  group_by(year) %>% 
  summarize(mean_max = mean(tmax, na.rm = TRUE), 
            mean_min = mean(tmin, na.rm = TRUE))

a = ggplot(fl_annual,
           aes(x=year, mean_max))+
  geom_point(col="orangered", alpha = 0.5)+
  scale_y_continuous(limits=c(min(fl_annual$mean_min),
                              max(fl_annual$mean_max)))+
geom_point(data=fl_annual, aes(x=year, mean_min), col="lightseagreen", alpha = 0.5)

a

a =a+stat_smooth(method="lm", col="orangered")
a

```

```{r, annual final graph, include = TRUE}

a+ 
  stat_smooth(data=fl_annual,
               aes(x=year,mean_min),
               col="light seagreen", method="lm")+
  theme_minimal()+
  labs(x = "Year", 
       y = "Average Annual Temperature (ºF)", 
       title = "Average daily minimmum and maximum average temperatures from 1999-2020", 
       caption = "Here the average daily minimum and maximum temperatures are shown over time. \nThe red dots represent the average maximum daily temperature, and the red line is a linear trendline\n of the changes over time. The blue dots represent the average minimum daily temperature, and the blue line is a\n linear trensline representing changes over time.")
```


```{r, include = FALSE}
library(Kendall)

MannKendall(florida$tmin)
#tau = 0.106, 2-sided pvalue =< 2.22e-16

MannKendall(florida$tmax)
#tau = 0.122, 2-sided pvalue =< 2.22e-16


t.test(subset(florida$tmin, florida$year %in% 1990:2009), subset(florida$tmin, florida$year %in% 2010:2020))


# data:  subset(florida$tmin, florida$year %in% 1990:2009) and subset(florida$tmin, florida$year %in% 2010:2020)
# t = -7.6443, df = 7602.3, p-value = 2.358e-14
# alternative hypothesis: true difference in means is not equal to 0
# 95 percent confidence interval:
#  -1.871952 -1.107829
# sample estimates:
# mean of x mean of y 
#  71.38860  72.87849 



```











### Discussion 
Why are these metrics likely to be relevant, given the climate change impacts you expect at this location?

Use some good old T, rank-sum, or other statistical tests,


### Citations 



